Assessing \hydrogen{} and \carbon{} Network Strategies in Europe

### 2023-12-06 report/submissions/journal/paper/paper.tex
shorten/find comprimize for 3.2

% Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model? An initial hypothesis may be that the cost assumptions for \carbon{}
%  and \hydrogen{} networks play a role. However, the amount of system costs spent on either network is substantially lower than the cost difference between the models (20 billion euros less total system costs in the \hydrogengrid{} model, while it spends 10 billion euros on the \hydrogen{} grid and the \carbongrid{} model spends 8 billion euros on the \carbon{} grid). Furthermore, we have conducted a sensitivity analysis with cost assumptions for the \carbon{} network reduced to 50\% of the default assumptions, and the results show that the \hydrogen{} network still outperforms the \carbon{} network in terms of total system costs.
To find the reason for why the \hydrogen{} network achieves lower system costs than the \carbon{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems. Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (upper row) and hydrogen (lower row) sectors for the \carbongrid{} (left) and \hydrogen{} (right) model. In each region, circles represent the production (top half-circle) and consumption (bottom half-circle) of the respective energy carrier, divided into shares of each technology and the area sizes correspond to production and consumption volumes. Differences between production and consumption in a given node correspond to import or export volumes through the connected networks, represented as lines with their size and arrows indicating volumes and direction of average flows. Each region's color represents the demand-weighted average price of the considered carrier in that region.

There are four notable dynamics or differences between the two grid strategies:
First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together. In the \carbongrid{} model, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions. In the \hydrogengrid{} model, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \carbon{} from point sources.
Second, the \carbongrid{} model captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy. Instead, the carbon grid transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen. The \hydrogengrid{} model directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
Third, sequestration in the \hydrogengrid{} model occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions. In the \carbongrid{} model, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \carbon{} from neighboring regions. However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices. Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load). Thus, for FT fuel synthesis sites in Northern Europe with more electricity from wind power are preferred to sites in Southern Europe, powered mostly by solar. Methanathion and Methanolisation both have higher flexibility, being able to ramp down to 50\% load during the night, for example. Thus, these fuel production sites prefer the best renewable resource sites in Southern Europe. And finally, as methanation produces synthetic methane that needs to be fed into the gas grid, methanation prefers those sites with good interconnection to the existing gas network. Since Italy possesses a large gas network with easy interconnection to Central Europe, methanation prefers sites in Southern Italy (and Greece, which would be connected to Southern Italy) over the Iberian Peninsula, where Methanolisation is more prevalent.

In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane). DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
In the absence of a \carbon{} network and with a \hydrogengrid{} on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it. The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration. Carbonaceous fuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from renewables, with some additional electrolysis and DAC required to fully meet the demand.
% Sequestered carbon is nearly always transported before

@@

Together with understanding the dynamics described above, we can highlight three cost advantages of the \hydrogen{} network in contrast to the \carbon{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC. However, the \hydrogengrid{} model produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR. Furthermore, without a \hydrogen{} network, additional \hydrogen{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
And finally, the \carbongrid{} model spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
% TODO: why??
The costs of the respective networks, on the other hand, are almost equal: the \carbongrid{} model spends 8 billion euros on the \carbon{} network, while the \hydrogengrid{} model even spends 10 billion euros on the \hydrogen{} network. These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \carbongrid{} and 765 billion euros for the \hydrogengrid{} model, see Fig.~\ref{fig:objective_heatmap}).

>>>>

% Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model? An initial hypothesis may be that the cost assumptions for \carbon{}
%  and \hydrogen{} networks play a role. However, the amount of system costs spent on either network is substantially lower than the cost difference between the models (20 billion euros less total system costs in the \hydrogengrid{} model, while it spends 10 billion euros on the \hydrogen{} grid and the \carbongrid{} model spends 8 billion euros on the \carbon{} grid). Furthermore, we have conducted a sensitivity analysis with cost assumptions for the \carbon{} network reduced to 50\% of the default assumptions, and the results show that the \hydrogen{} network still outperforms the \carbon{} network in terms of total system costs.
To find the reason for why the \hydrogen{} network achieves lower system costs than the \carbon{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems. Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (upper row) and hydrogen (lower row) sectors for the \carbongrid{} (left) and \hydrogen{} (right) model. Each map includes circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines representing the import/export volumes and flow directions. The color of each region corresponds to the average price of the carrier.

The \carbongrid{} model captures large amounts of low-cost carbon from bioenergy and industrial activities, especially in Central and Eastern Europe (top left), almost none of which is used locally. A large part is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used to supplement low-cost hydrogen in central plants for electrolysis, Fischer-Tropsch and methanolization (bottom left). A second part is transported to methanation plants such as those in Italy and Greece, from where the gas is fed into the gas grid, which also is set up for distribution from large gas import hubs. Finally, the third part is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is eventually enriched with \carbon{} from DAC to be finally sequestered. In absence of a \hydrogen{} network, the system distributes electrolysis where it is needed for industrial processes and fuel cell land transport and emphasizes gas imports into Central Europe.

The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions (lower right). The main transportation routes go from Spain and the United Kingdom to Central Europe. In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources, leading to a decentralized CU system across the continent. DAC and local emissions nearby the coast enable sequestration in the North Sea, Baltic Sea, Mediterranean Sea and, in particular, in Portugal (upper right).
% Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula. In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above). In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network). Surplus hydrogen from lower FT fuel production is exported to Central Europe.

The results illustrate the fundamentally different approaches of the grids: a \carbon{} grid prefers a decentralized CC system and a co-location of electrolysis and CU centrally in regions with abundant renewable resources; a \hydrogen{} grid prefers a centralized electrolysis system and a decentralized CC and CU system distributed jointly across the continent. A \hydrogen{} network not only unlocks the potential of carbon capture from point sources, but also reduces the energy shortages in energy-intensive areas.

% There are four notable dynamics or differences between the two grid strategies:
% First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together. In the \carbongrid{} model, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions. In the \hydrogengrid{} model, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \carbon{} from point sources.
% Second, the \carbongrid{} model captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy. Instead, the carbon grid transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen. The \hydrogengrid{} model directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
% Third, sequestration in the \hydrogengrid{} model occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions. In the \carbongrid{} model, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \carbon{} from neighboring regions. However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
% And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices. Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load). Thus, for FT fuel synthesis sites in Northern Europe with more electricity from wind power are preferred to sites in Southern Europe, powered mostly by solar. Methanathion and Methanolisation both have higher flexibility, being able to ramp down to 50\% load during the night, for example. Thus, these fuel production sites prefer the best renewable resource sites in Southern Europe. And finally, as methanation produces synthetic methane that needs to be fed into the gas grid, methanation prefers those sites with good interconnection to the existing gas network. Since Italy possesses a large gas network with easy interconnection to Central Europe, methanation prefers sites in Southern Italy (and Greece, which would be connected to Southern Italy) over the Iberian Peninsula, where Methanolisation is more prevalent.

% In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane). DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
% In the absence of a \carbon{} network and with a \hydrogengrid{} on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it. The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration. Carbonaceous fuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from renewables, with some additional electrolysis and DAC required to fully meet the demand.
% Sequestered carbon is nearly always transported before

@@

% Together with understanding the dynamics described above, we can highlight three cost advantages of the \hydrogen{} network in contrast to the \carbon{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC. However, the \hydrogengrid{} model produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR. Furthermore, without a \hydrogen{} network, additional \hydrogen{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
% And finally, the \carbongrid{} model spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
% TODO: why??
% The costs of the respective networks, on the other hand, are almost equal: the \carbongrid{} model spends 8 billion euros on the \carbon{} network, while the \hydrogengrid{} model even spends 10 billion euros on the \hydrogen{} network. These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \carbongrid{} and 765 billion euros for the \hydrogengrid{} model, see Fig.~\ref{fig:objective_heatmap}).
